### 2025-12-01 Paper Draft.Rmd
Initial commit.

---
title: "Predictive Modeling of Drinking Water Treatment Process Chemical Doses"
author: "Dawson Carney, Reece Carmody, Ethan Schilling, Joshua Tobey"
date: "5 December 2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction