### 2025-12-04 Exploratory Analysis.Rmd
Tried clustering

---
title: "Exploratory Analysis"
author: "Reece Carmody"
date: "2025-11-17"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)

data = read.csv("RWAR_turb_cond_counts_pH_temp.csv")

data$Timestamp = mdy_hm(data$Timestamp)

# Trying to determine what rows have outliers.
# The goal is to go by column, remove outliers & average that day's data, then ultimately combine every column after it's cleaned to create our fully cleaned data frame.
# Then combine with alkalinity and hardness & coagulant doses data??

ggplot(data, aes(x = Timestamp, y = Turbidity..NTU)) +
  geom_point()

data[data$Turbidity..NTU > 10, ] # row 2701 and 5446

ggplot(data, aes(x = Timestamp, y = Conductivity..mS.cm.1)) +
  geom_point()

data[data$Conductivity..mS.cm.1 > 15, ] # row 5270

# Not sure what to do with particle count (PC) data just yet

ggplot(data, aes(x = Timestamp, y = pH..SU)) +
  geom_point()

ggplot(data, aes(x = Timestamp, y = Temperature..deg.C)) +
  geom_point()

```

```{r}

df1 = read.csv("RWAR_temp_pH_cond_turb_cleaned.csv")
df2 = read.csv("alkalinity_hardness_cleaned.csv")
df3 = read.csv("CoagulantDoses_2017-20.csv")

df1 = df1 %>%
  mutate(Date = ymd(Timestamp))
df2 = df2 %>%
  mutate(Date = mdy(Date))
df3 = df3 %>%
  mutate(Date = mdy(Date))

# This merge is only keeping the rows that are in both data frames
# Not using df3 because I don't want our target variables to be included in the clustering
merged = df1 %>% inner_join(df2, by = "Date")
merged$Date = NULL # We already have the Timestamp column

cluster_data = merged %>%
  select(where(is.numeric)) %>%
  drop_na()

cluster_data = scale(cluster_data) # Standardize the data

cluster_fit = kmeans(cluster_data, centers = 5, nstart = 25)

merged$cluster = cluster_fit$cluster

# Interpretation

table(merged$cluster)
cluster_fit$centers

ggplot(merged, aes(x = cluster_data[ , "mean_Temp"], y = cluster_data[ , "mean_pH"], # you can compare any two columns
                   color = factor(cluster))) +
  geom_point() +
  labs(title = "K-Means Clustering", color = "Cluster") +
  theme_bw()

```