### 2025-12-04 Exploratory_Analysis_Josh_Tobey.Rmd
WIP before rebase

---
title: "Exploratory_Analysis_Josh_Tobey"
author: "Joshua Tobey"
date: "2025-11-17"
output: html_document
---
```{r}
install.packages(c("usethis", "gitcreds"))
```

```{r}
library(tidyverse)
library(tidymodels)
library(dplyr)
library(knitr)

```

```{r}
exploratory_df <- read_csv("RWAR_DOSING_clean.csv")
colnames(exploratory_df)
```

```{r}
ggplot(exploratory_df) +
  geom_point(aes(x = `Turbidity, NTU`,
                 y = `Temperature, deg C`,
                 color = `Temperature, deg C`)) +
  scale_color_gradient(low = 'yellow', high = 'green')
```

```{r}
head(exploratory_df)

```
```{r}
exploratory_df$Timestamp = mdy_hm(exploratory_df$Timestamp)
ggplot(exploratory_df, aes(x = Timestamp,
                           y = `pH, SU`,
                           color = `pH, SU`)) +
  geom_line() + 
  scale_color_gradient(low = 'blue', high = 'red')

```

```{r}
unique(exploratory_df$`Conductivity, mS cm-1`)
```
```{r}
ggplot(exploratory_df) +
  geom_jitter(aes(x = `Turbidity, NTU`,
                 y = `Conductivity, mS cm-1`,
                 color = `Conductivity, mS cm-1`), width = 1, height = 3) +
  scale_color_gradient(low = 'skyblue', high = 'orange')

```
Nothing of real interest here.
```{r}
coag_dose_df <- read_csv("CoagulantDoses_2017-20.csv")
head(coag_dose_df)

```
```{r}
#exploratory_df <- exploratory_df %>%
  #mutate(`dosing (exploratory)` = coag_dose_df$`Cationic Polimer, mg/L` )
length(exploratory_df) == length(coag_dose_df)
```
```{r}

coag_dose_df$Date <- mdy(coag_dose_df$Date)
ggplot(coag_dose_df) +
  geom_line(aes(x = Date,
                y = `Cationic Polimer, mg/L`))

```
```{r}
alk_df <- read_csv("alkalinity_hardness_cleaned.csv")
head(alk_df)

```
```{r}
alk_df$Date <- mdy(alk_df$Date)
ggplot(alk_df) +
  geom_line(aes(x = Date,
                y = alkalinity),
            color = 'orange') +
  geom_line(aes(x = Date,
                y = hardness),
            color = 'blue')

```
```{r}
generic_mlr <- linear_reg() |>
  set_mode("regression") |>
  set_engine("lm")

mlr_recipe <- recipe(alkalinity ~ hardness, data = alk_df)
mlr_wf <- workflow() |> 
  add_model(generic_mlr) |> 
  add_recipe(mlr_recipe)
fit(mlr_wf, data = alk_df)
```

```{r}
time_lag_1 <- 
time_lag_2 <- 
time_lag_3 <-

weekly_log_spec <- logistic_reg()
weekly_log_fit <- weekly_log_spec |>
  fit(Direction ~ Lag1 + 
      Lag2, 
      family = "binomial", 
      data = Weekly)
weekly_log_fit |>
  pluck("fit") |>
  summary()

```

```{r}
cross_validation <- initial_split(exploratory_df, prop = 0.6)
exploratory_train <- training(cross_validation)
exploratory_test <- testing(cross_validation)
#OPTION 1:
linear_recipe <- recipe(`pH SU` ~ `Temperature deg C`, data = exploratory_df)
quad_recipe <- linear_recipe |> step_mutate(`Temperature deg C_2` = `Temperature deg C`^2)
cubic_recipe <- quad_recipe |> step_mutate(`Temperature deg C_3` = `Temperature deg C`^3)
quart_recipe <- cubic_recipe |> step_mutate(`Temperature deg C_4` = `Temperature deg C`^4)
#OPTION 2:
linear_recipe <- recipe(`pH SU` ~ ., data = exploratory_df)
quad_recipe <- linear_recipe |> step_poly(all_numeric_predictors(), degree = 2)
cubic_recipe <- quad_recipe |> step_poly(all_numeric_predictors(), degree = 3)
quart_recipe <- cubic_recipe |> step_poly(all_numeric_predictors(), degree = 4)

loocv_df <- loo_cv(dfi)
test_errors <- data.frame()

retrieve_test_error <- function(least_sqrs_model, testing_set) {
  augment(least_sqrs_model, new_data = testing_set) |>
    mutate(resid2 = .resid^2) |> 
    summarise(rmse = sqrt(mean(resid2))) |>
    pull(rmse)
}

for (i in seq_len(nrow(loocv_df))) {
  train_data = training(loocv_df[i,]$splits[[1]])
  test_data = testing(loocv_df[i,]$splits[[1]])

  Y1 <- workflow() |> add_model(lstSqrs_spec) |> add_recipe(linear_recipe) |> fit(train_data)
  Y2 <- workflow() |> add_model(lstSqrs_spec) |> add_recipe(quad_recipe) |> fit(train_data)
  Y3 <- workflow() |> add_model(lstSqrs_spec) |> add_recipe(cubic_recipe) |> fit(train_data)
  Y4 <- workflow() |> add_model(lstSqrs_spec) |> add_recipe(quart_recipe) |> fit(train_data)

  data.frame(model = c("linear", "quadratic", "cubic", "quartic"),
             rmse = c(retrieve_test_error(Y1, test_data), 
                   retrieve_test_error(Y2, test_data), 
                   retrieve_test_error(Y3, test_data), 
                   retrieve_test_error(Y4, test_data)),
             i = i) |>
    rbind(test_errors) -> test_errors

}
test_errors |> 
  group_by(model) |>
  summarise(rmse = sqrt(mean(rmse))) |>
  kable()

```