### 2025-12-04 Presentation DSCI445.Rmd
Updates to the Presentation

---
title: "DSCI 445 Presentation"
author: "Ethan Schilling, Joshua Tobey, Dawson Carney, Reece Carmody"
date: "2025-12-04"
output: beamer_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
library(patchwork)
```

## Slide 1

```{r}
TOC <- readRDS("TOC_plot.rds")
TDS <- readRDS("TDS_plot.rds")
TOC + TDS

```