### 2025-12-04 Exploratory Analysis Ethan Schilling.Rmd

---
title: "Exploratory Analysis Ethan Schilling"
output: html_document
---

```{r}
library(tidyverse)
library(knitr)
library(tidymodels)
library(dplyr)
```

```{r}

TDS_unfilt <- read_csv("RWAR_TDS.csv")
TOC_unfilt <- read_csv("RWAR_TOC.csv")

df_TDS <- TDS_unfilt %>%
  rename(date = 1, value = 2) %>%
  mutate(date = as.Date(date, format = "%d-%b-%y"))

TDS_plot <- ggplot(df_TDS, aes(x = date, y = value)) +
  geom_line() +
  labs(x = "Date", y = "TDS Value", title = "TDS over Time")

saveRDS(TDS_plot,"TDS_plot.rds")

df_TOC <- TOC_unfilt %>% 
  rename(date = 1, value = 2) %>%
  mutate(date = as.Date(date, format = "%d-%b-%y"))

TOC_plot <- ggplot(df_TOC, aes(x = date, y = value)) +
  geom_line() +
  labs(x = "Date", y = "TOC Value", title = "TOC over Time")

saveRDS(TOC_plot,"TOC_plot.rds")

```