### 2025-12-08 Exploratory Analysis.Rmd

---
title: "Exploratory Analysis"
author: "Reece Carmody"
date: "2025-11-17"
output: pdf_document
---

```{r}
print("GitHub test")

```