### 2025-12-09 Tree Models Reece C.Rmd
Decision tree & random forest

---
title: "Tree Models"
output: pdf_document
---

```{r}
set.seed(445)

library(dplyr)
library(tidyverse)
library(tidymodels)

# Decision Tree

# For now just predicting numerical amounts, not classifying
data = read.csv("RWAR_dosing_lag_incdec.csv")
data = as.data.frame(data)
data$FeCl_change = NULL
data$Polymer_change = NULL

# Set up training & testing data

# Can change these if you want to modify the tree/data
tree_depth = 5 # tree depth
min_n = 50 # minimum observations needed to further split
train_proportion = 0.8 # split of data

n = nrow(data)
indices = sample(1:n, size = floor(train_proportion * n))

train_data = data[indices, ]
test_data  = data[-indices, ]

# Create the tree

recipe = recipe(FeCl_dose ~ ., data = train_data) %>% # Change target variable here
  step_rm(Timestamp) %>%
  step_rm(Polymer_dose) # Remove the other target variable here

tree_spec = decision_tree(tree_depth = tree_depth, min_n = min_n) %>%
  set_engine("rpart") %>%
  set_mode("regression")

workflow = workflow() %>%
  add_recipe(recipe) %>%
  add_model(tree_spec)

fit_tree = workflow %>%
  fit(data = train_data)

# Assess accuracy

predictions = fit_tree %>%
  predict(test_data) %>%
  bind_cols(test_data)

actual = predictions$`FeCl_dose` # Change target variable here
predicted = predictions$.pred

rmse = sqrt(mean((actual - predicted)^2))
print(paste("RMSE: ", rmse))

r_squared = cor(actual, predicted)^2
print(paste("R Squared: ", r_squared))

# Plot the tree

library(rpart.plot)

fit_tree %>%
  extract_fit_engine() %>%
  rpart.plot(roundint = FALSE)

```

```{r}
set.seed(445)

library(dplyr)
library(tidyverse)
library(tidymodels)

# Random Forest

# For now just predicting numerical amounts, not classifying
data = read.csv("RWAR_dosing_lag_incdec.csv")
data = as.data.frame(data)
data$FeCl_change = NULL
data$Polymer_change = NULL

# Set up training & testing data

# Can change these if you want to modify the tree/data
trees = 500 # number of trees in the forest
min_n = 50 # minimum observations needed to further split
train_proportion = 0.8 # split of data

n = nrow(data)
indices = sample(1:n, size = floor(train_proportion * n))

train_data = data[indices, ]
test_data  = data[-indices, ]

# Create the tree

recipe = recipe(FeCl_dose ~ ., data = train_data) %>% # Change target variable here
  step_rm(Timestamp) %>%
  step_rm(Polymer_dose) # Remove the other target variable here

rf_spec = rand_forest(trees = trees, min_n = min_n) %>%
  set_engine("ranger") %>%
  set_mode("regression")

workflow = workflow() %>%
  add_recipe(recipe) %>%
  add_model(rf_spec)

fit_rf = workflow %>%
  fit(data = train_data)

# Assess accuracy

predictions = fit_rf %>%
  predict(test_data) %>%
  bind_cols(test_data)

actual = predictions$`FeCl_dose` # Change target variable here
predicted = predictions$.pred

rmse = sqrt(mean((actual - predicted)^2))
print(paste("RMSE: ", rmse))

r_squared = cor(actual, predicted)^2
print(paste("R Squared: ", r_squared))

```