### 2025-12-09 Exploratory_Analysis_Josh_Tobey.Rmd
Preliminary fitting of GAM.

---
title: "Exploratory_Analysis_Josh_Tobey"
author: "Joshua Tobey"
date: "2025-11-17"
output: html_document
---
```{r}
install.packages(c("usethis", "gitcreds"))
```

```{r}
library(tidyverse)
library(tidymodels)
library(dplyr)
library(knitr)
library(lubridate)

```

```{r}
exploratory_df <- read_csv("RWAR_dosing_cleaned.csv")
colnames(exploratory_df)
```

```{r}
ggplot(exploratory_df) +
  geom_point(aes(x = mean_Turbidity,
                 y = mean_Temp,
                 color = mean_Temp)) +
  scale_color_gradient(low = 'yellow', high = 'green')
```

```{r}
head(exploratory_df)

```

```{r}
exploratory_df$Timestamp = ymd(exploratory_df$Timestamp)
exploratory_df <- exploratory_df[order(exploratory_df$Timestamp), ]
ggplot(exploratory_df, aes(x = Timestamp,
                           y = mean_pH,
                           color = mean_pH)) +
  geom_line() + 
  scale_color_gradient(low = 'blue', high = 'red')

```

```{r}
library(patchwork)
plot1 <- ggplot(exploratory_df, aes(x = Timestamp,
                           y = mean_pH,
                           color = mean_pH)) +
  geom_line() + 
  scale_color_gradient(low = 'blue', high = 'red') +
  theme_minimal() +
  ggtitle("Average pH over Time")

plot2 <- ggplot(exploratory_df, aes(x = Timestamp,
                           y = Cationic.Polimer..mg.L,
                           color = Cationic.Polimer..mg.L)) +
  geom_line() + 
  scale_color_gradient(low = 'yellow', high = 'green') +
  theme_minimal() +
  ggtitle("Polymer Coagulant over Time")

plot1/plot2
```

```{r}
ggplot(exploratory_df) +
  geom_jitter(aes(x = mean_Turbidity,
                 y = mean_Conductivity,
                 color = mean_Conductivity), width = 1, height = 3) +
  scale_color_gradient(low = 'skyblue', high = 'orange')

```
Nothing of real interest here.
```{r}
coag_dose_df <- read_csv("CoagulantDoses_2017-20.csv")
head(coag_dose_df)

```
```{r}
#exploratory_df <- exploratory_df %>%
  #mutate(`dosing (exploratory)` = coag_dose_df$`Cationic Polimer, mg/L` )
length(exploratory_df) == length(coag_dose_df)
```
```{r}

coag_dose_df$Date <- mdy(coag_dose_df$Date)
ggplot(coag_dose_df) +
  geom_line(aes(x = Date,
                y = `Cationic Polimer, mg/L`))

```
```{r}
alk_df <- read_csv("alkalinity_hardness_cleaned.csv")
head(alk_df)

```
```{r}
alk_df$Date <- mdy(alk_df$Date)
ggplot(alk_df) +
  geom_line(aes(x = Date,
                y = alkalinity),
            color = 'orange') +
  geom_line(aes(x = Date,
                y = hardness),
            color = 'blue')

```
```{r}
generic_mlr <- linear_reg() |>
  set_mode("regression") |>
  set_engine("lm")

mlr_recipe <- recipe(alkalinity ~ hardness, data = alk_df)
mlr_wf <- workflow() |> 
  add_model(generic_mlr) |> 
  add_recipe(mlr_recipe)
fit(mlr_wf, data = alk_df)
```

```{r}
weekly_log_spec <- logistic_reg()
weekly_log_fit <- weekly_log_spec |>
  fit(Direction ~ Lag1 + 
      Lag2, 
      family = "binomial", 
      data = Weekly)
weekly_log_fit |>
  pluck("fit") |>
  summary()

```

```{r}
cross_validation <- initial_split(exploratory_df, prop = 0.6)
exploratory_train <- training(cross_validation)
exploratory_test <- testing(cross_validation)
#OPTION 1:
linear_recipe <- recipe(`pH SU` ~ `Temperature deg C`, data = exploratory_df)
quad_recipe <- linear_recipe |> step_mutate(`Temperature deg C_2` = `Temperature deg C`^2)
cubic_recipe <- quad_recipe |> step_mutate(`Temperature deg C_3` = `Temperature deg C`^3)
quart_recipe <- cubic_recipe |> step_mutate(`Temperature deg C_4` = `Temperature deg C`^4)
#OPTION 2:
linear_recipe <- recipe(`pH SU` ~ ., data = exploratory_df)
quad_recipe <- linear_recipe |> step_poly(all_numeric_predictors(), degree = 2)
cubic_recipe <- quad_recipe |> step_poly(all_numeric_predictors(), degree = 3)
quart_recipe <- cubic_recipe |> step_poly(all_numeric_predictors(), degree = 4)

loocv_df <- loo_cv(dfi)
test_errors <- data.frame()

retrieve_test_error <- function(least_sqrs_model, testing_set) {
  augment(least_sqrs_model, new_data = testing_set) |>
    mutate(resid2 = .resid^2) |> 
    summarise(rmse = sqrt(mean(resid2))) |>
    pull(rmse)
}

for (i in seq_len(nrow(loocv_df))) {
  train_data = training(loocv_df[i,]$splits[[1]])
  test_data = testing(loocv_df[i,]$splits[[1]])

  Y1 <- workflow() |> add_model(lstSqrs_spec) |> add_recipe(linear_recipe) |> fit(train_data)
  Y2 <- workflow() |> add_model(lstSqrs_spec) |> add_recipe(quad_recipe) |> fit(train_data)
  Y3 <- workflow() |> add_model(lstSqrs_spec) |> add_recipe(cubic_recipe) |> fit(train_data)
  Y4 <- workflow() |> add_model(lstSqrs_spec) |> add_recipe(quart_recipe) |> fit(train_data)

  data.frame(model = c("linear", "quadratic", "cubic", "quartic"),
             rmse = c(retrieve_test_error(Y1, test_data), 
                   retrieve_test_error(Y2, test_data), 
                   retrieve_test_error(Y3, test_data), 
                   retrieve_test_error(Y4, test_data)),
             i = i) |>
    rbind(test_errors) -> test_errors

}
test_errors |> 
  group_by(model) |>
  summarise(rmse = sqrt(mean(rmse))) |>
  kable()

```

```{r}
ggplot(exploratory_df) +
  geom_point(aes(x = mean_pH,
                 y = mean_Temp))
```
```{r}
RWAR_dosing_lag_incdec <- read_csv("RWAR_dosing_lag_incdec.csv")
head(RWAR_dosing_lag_incdec)

```
```{r}
RWAR_dosing_lag_incdec <- RWAR_dosing_lag_incdec %>%
  select(-c(FeCl_change, Polymer_change, Timestamp))
head(RWAR_dosing_lag_incdec)
```

```{r}
library(mgcv)
library(parsnip)

gam_spec <- gen_additive_mod() |> 
  set_engine("mgcv") |> 
  set_mode("regression")
gam_rec <- recipe(FeCl_dose ~ Polymer_dose + mean_Temp + mean_pH + mean_Conductivity +
                 mean_Turbidity + Temp_lag1 + Temp_lag2 +
                 pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2 +
                 Turb_lag1 + Turb_lag2, data = RWAR_dosing_lag_incdec)

gam_wf <- workflow() |>
  add_model(gam_spec) |>
  add_recipe(gam_rec)

gam_fit <- gam(
  FeCl_dose ~ s(Polymer_dose) + s(mean_Temp) + s(mean_pH) + s(mean_Conductivity) +
               s(mean_Turbidity) + s(Temp_lag1) + s(Temp_lag2) +
               s(pH_lag1) + s(pH_lag2) + s(Cond_lag1) + s(Cond_lag2) +
               s(Turb_lag1) + s(Turb_lag2),
  data = RWAR_dosing_lag_incdec
)

gam_fit |> tidy()
gam_fit |> glance()
```
```{r}

dat_grid_dose <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(Polymer_dose = seq(min(RWAR_dosing_lag_incdec$Polymer_dose),
                                      max(RWAR_dosing_lag_incdec$Polymer_dose),
                                      length.out = 100))
dat_grid_temp <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(mean_Temp = seq(min(RWAR_dosing_lag_incdec$mean_Temp), 
                         max(RWAR_dosing_lag_incdec$mean_Temp),
                                      length.out = 100))
dat_grid_pH <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(mean_pH = seq(min(RWAR_dosing_lag_incdec$mean_pH), max(RWAR_dosing_lag_incdec$mean_pH),
                                      length.out = 100))

dat_grid_cond <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(mean_Conductivity = seq(min(RWAR_dosing_lag_incdec$mean_Conductivity), max(RWAR_dosing_lag_incdec$mean_Conductivity),
                                      length.out = 100))
dat_grid_turb <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(mean_Turbidity = seq(min(RWAR_dosing_lag_incdec$mean_Turbidity), max(RWAR_dosing_lag_incdec$mean_Turbidity),
                                      length.out = 100))
dat_grid_temp_lag1 <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(Temp_lag1 = seq(min(RWAR_dosing_lag_incdec$Temp_lag1), max(RWAR_dosing_lag_incdec$Temp_lag1),
                                      length.out = 100))
dat_grid_temp_lag2 <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(Temp_lag2 = seq(min(RWAR_dosing_lag_incdec$Temp_lag2), max(RWAR_dosing_lag_incdec$Temp_lag2),
                                      length.out = 100))
dat_grid_pH_lag1 <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(pH_lag1 = seq(min(RWAR_dosing_lag_incdec$pH_lag1), max(RWAR_dosing_lag_incdec$pH_lag1),
                                      length.out = 100))
dat_grid_pH_lag2 <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(pH_lag2 = seq(min(RWAR_dosing_lag_incdec$pH_lag2), max(RWAR_dosing_lag_incdec$pH_lag2),
                                      length.out = 100))
dat_grid_cond_lag1 <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(Cond_lag1 = seq(min(RWAR_dosing_lag_incdec$Cond_lag1), max(RWAR_dosing_lag_incdec$Cond_lag1),
                                      length.out = 100))
dat_grid_cond_lag2 <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(Cond_lag2 = seq(min(RWAR_dosing_lag_incdec$Cond_lag2), max(RWAR_dosing_lag_incdec$Cond_lag2),
                                      length.out = 100))
dat_grid_turb_lag1 <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(Turb_lag1 = seq(min(RWAR_dosing_lag_incdec$Turb_lag1), max(RWAR_dosing_lag_incdec$Turb_lag1),
                                      length.out = 100))
dat_grid_turb_lag2 <- RWAR_dosing_lag_incdec |> 
  summarize(
    across(everything(), mean, na.rm = TRUE)
  ) |> 
  slice(rep(1, 100)) |> 
  mutate(Turb_lag2 = seq(min(RWAR_dosing_lag_incdec$Turb_lag2), max(RWAR_dosing_lag_incdec$Turb_lag2),
                                      length.out = 100))

pred1 <- predict(gam_fit, newdata = dat_grid_dose, type = "response", se.fit = TRUE)
dose_pred <- dat_grid_dose |> mutate(
  .fitted = pred1$fit,
  .lower = pred1$fit - 2 * pred1$se.fit,
  .upper = pred1$fit + 2 * pred1$se.fit
  )
pred2 <- predict(gam_fit, newdata = dat_grid_temp, type = "response", se.fit = TRUE)
temp_pred <- dat_grid_temp |> mutate(
  .fitted = pred2$fit,
  .lower = pred2$fit - 2 * pred2$se.fit,
  .upper = pred2$fit + 2 * pred2$se.fit
)
pH_pred <- augment(gam_fit, new_data = dat_grid_pH)
cond_pred <- augment(gam_fit, new_data = dat_grid_cond)
turb_pred <- augment(gam_fit, new_data = dat_grid_turb)
tempLag1_pred <- augment(gam_fit, new_data = dat_grid_temp_lag1)
tempLag2_pred <- augment(gam_fit, new_data = dat_grid_temp_lag2)
pHLag1_pred <- augment(gam_fit, new_data = dat_grid_pH_lag1)
pHLag2_pred <- augment(gam_fit, new_data = dat_grid_pH_lag2)
condLag1_pred <- augment(gam_fit, new_data = dat_grid_cond_lag1)
condLag2_pred <- augment(gam_fit, new_data = dat_grid_cond_lag2)
turbLag1_pred <- augment(gam_fit, new_data = dat_grid_turb_lag1)
turbLag2_pred <- augment(gam_fit, new_data = dat_grid_turb_lag2)

```

```{r}
ggplot(dose_pred, aes(Polymer_dose, .fitted)) +
  geom_line() + 
  geom_ribbon(aes(ymin = .lower, ymax = .upper), alpha = 0.2)
```