### 2025-12-10 Tree Models Reece C.Rmd

---
title: "Tree Models"
output: pdf_document
---

```{r}
set.seed(445)

library(dplyr)
library(tidyverse)
library(tidymodels)
library(rpart.plot)
library(knitr)
library(kableExtra)

# Decision Tree

# For now just predicting numerical amounts, not classifying
data = read.csv("RWAR_dosing_lag_incdec.csv")
data = as.data.frame(data)
data$FeCl_change = NULL
data$Polymer_change = NULL

# Set up training & testing data

# Can change these if you want to modify the tree/data
tree_depth = 5 # tree depth
min_n = 50 # minimum observations needed to further split

train_data = data %>%
  filter(year(Timestamp) <= 2019)
test_data = data %>%
  filter(year(Timestamp) >= 2020)

# Create the tree

recipe = recipe(FeCl_dose ~ ., data = train_data) %>% # Change target variable here
  step_rm(Timestamp) %>%
  step_rm(Polymer_dose) # Remove the other target variable here

tree_spec = decision_tree(tree_depth = tree_depth, min_n = min_n) %>%
  set_engine("rpart") %>%
  set_mode("regression")

workflow = workflow() %>%
  add_recipe(recipe) %>%
  add_model(tree_spec)

fit_tree = workflow %>%
  fit(data = train_data)

# Assess accuracy

predictions = fit_tree %>%
  predict(test_data) %>%
  bind_cols(test_data)

actual = predictions$`FeCl_dose` # Change target variable here
predicted = predictions$.pred

rmse = sqrt(mean((actual - predicted)^2))
r_squared = cor(actual, predicted)^2
mae = mean(abs(actual-predicted))

metrics = data.frame(
  Metric = c("RMSE", "R-Squared", "MAE"),
  Value  = c(rmse, r_squared, mae)
)

kable(metrics, booktabs = TRUE, digits = 3) %>%
  kable_styling(position = "center", font_size = 12)

# Plot the tree

fit_tree %>%
  extract_fit_engine() %>%
  rpart.plot(roundint = FALSE)

```

```{r}
set.seed(445)

library(dplyr)
library(tidyverse)
library(tidymodels)
#library(yardstick)
library(vip)

# Random Forest

# For now just predicting numerical amounts, not classifying
data = read.csv("RWAR_dosing_lag_incdec.csv")
data = as.data.frame(data)
data$FeCl_change = NULL
data$Polymer_change = NULL

# Set up training & testing data

# Can change these if you want to modify the tree/data
trees = 500 # number of trees in the forest
min_n = 50 # minimum observations needed to further split

train_data = data %>%
  filter(year(Timestamp) <= 2019)
test_data = data %>%
  filter(year(Timestamp) >= 2020)

# Create the tree

recipe = recipe(FeCl_dose ~ ., data = train_data) %>% # Change target variable here
  step_rm(Timestamp) %>%
  step_rm(Polymer_dose) # Remove the other target variable here

rf_spec = rand_forest(trees = trees, min_n = min_n) %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("regression")

workflow = workflow() %>%
  add_recipe(recipe) %>%
  add_model(rf_spec)

fit_rf = workflow %>%
  fit(data = train_data)

# Assess accuracy

predictions = fit_rf %>%
  predict(test_data) %>%
  bind_cols(test_data)

actual = predictions$`FeCl_dose` # Change target variable here
predicted = predictions$.pred

# Metrics and plots

model_metrics = predictions %>%
  metrics(truth = FeCl_dose, estimate = .pred) # Change target variable here

print(model_metrics)

ggplot(predictions, aes(x = FeCl_dose, y = .pred)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1) +
  labs(title = "Random Forest Accuracy",
       x = "Actual",
       y = "Predicted") +
  theme_bw()

vip(fit_rf)

```