### 2025-12-10 Exploratory Analysis Ethan SchillingREAL.Rmd

---
title: "Exploratory Analysis Ethan Schilling"
output: html_document
---

```{r}
library(tidyverse)
library(knitr)
library(tidymodels)
library(dplyr)
library(Metrics)
library(kableExtra)

```

```{r}

TDS_unfilt <- read_csv("~/DSCI445/project-4/RWAR_TDS.csv")
TOC_unfilt <- read_csv("~/DSCI445/project-4/RWAR_TOC.csv")
DOC_unfilt <- read_csv("~/DSCI445/project-4/RWAR_DOC.csv")

df_TDS <- TDS_unfilt %>%
  rename(date = 1, value = 2) %>%
  mutate(date = as.Date(date, format = "%d-%b-%y"))

TDS_plot <- ggplot(df_TDS, aes(x = date, y = value)) +
  geom_line() +
  labs(x = "Date", y = "Total Dissolved Solids mg/L", title = "Total Dissolved Solids over Time")

df_TOC <- TOC_unfilt %>% 
  rename(date = 1, value = 2) %>%
  mutate(date = as.Date(date, format = "%d-%b-%y"))

TOC_plot <- ggplot(df_TOC, aes(x = date, y = value)) +
  geom_line() +
  labs(x = "Date", y = "Total Organinc Carbon mg/L", title = "Total Organic Carbon over Time")

df_DOC <- DOC_unfilt %>%
  rename(date = 1, value = 2) %>%
  mutate(date = as.Date(date, format = "%d-%b-%y"))

DOC_plot <- ggplot(df_DOC, aes(x = date, y = value)) +
  geom_line() +
  labs(x = "Date", y = "Dissolved Organic Carbon mg/L", title = "Dissolved Organic Carbon over Time")

#saveRDS(TDS_plot,"TDS_plot.rds")
#saveRDS(TOC_plot,"TOC_plot.rds")
#saveRDS(DOC_plot,"DOC_plot.rds")

TDS_plot
TOC_plot
DOC_plot

```

```{r}
library(lubridate)
library(dplyr)
library(ggplot2)
library(tidyr)
library(Metrics)
library(kableExtra)
library(gt)

dosing_data <- read.csv("~/DSCI445/project-4/RWAR_dosing_cleaned.csv")

dosing_long <- dosing_data %>%
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "Characteristics",
    values_to = "Value"
  )

ggplot(dosing_long, aes(x = Timestamp, y = Value, color = Ferric.chloride..mg.L)) +
  geom_point(size = 0.7) +
  facet_wrap(~Characteristics, scales = "free_y") +
  scale_color_viridis_c(option = "plasma") +
  theme_minimal() +
  labs(
    x = "Date",
    y = "Value",
    color = "Ferric Chloride (mg/L)",
    title = "Water Quality Characteristics Over Time by Dosing"
  )

RWAR_dosing <- read.csv("~/DSCI445/project-4/RWAR_dosing_lag_incdec.csv")

RWAR_dosing

RWAR_dosing <- RWAR_dosing %>% 
  mutate(Timestamp = ymd(Timestamp),
         Year = year(Timestamp))

RWAR_dosing

train <- RWAR_dosing %>% filter(Year %in% c(2018, 2019))
test  <- RWAR_dosing %>% filter(Year == 2020)

# All predictors

model0 <- lm(FeCl_dose ~ mean_Temp + mean_pH + mean_Conductivity + mean_Turbidity + Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2 + Turb_lag1 + Turb_lag2, data = train)

test$pred_model0 <- predict(model0, newdata = test)

# 1 Interaction Term

model1 = lm(FeCl_dose ~ mean_Temp * mean_pH + mean_Turbidity + Temp_lag1  + Temp_lag2 + Turb_lag1 +     Turb_lag2,data = train)
summary(model1)

test$pred_model1 <- predict(model1, newdata = test)

# 2 Interaction Terms

model2 = lm(FeCl_dose ~ mean_Temp * mean_pH + mean_Conductivity * mean_Turbidity +
                Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2, data = train)

test$pred_model2 <- predict(model2, newdata = test)

model3 = lm(FeCl_dose ~ mean_Temp * mean_Conductivity + Temp_lag1 + Temp_lag2 + Cond_lag1 + Cond_lag2,                 data = train)

test$pred_model3 <- predict(model3, newdata = test)

# 2 Interactions: Cond and PH and Temp and Cond

model4 = lm(FeCl_dose ~ mean_Conductivity * mean_pH + mean_Temp * mean_Conductivity +
                Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2,
              data = train)

test$pred_model4 <- predict(model4, newdata = test)

model5 = lm(FeCl_dose ~ mean_Temp + mean_pH + Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2,
              data = train)

test$pred_model5 <- predict(model5, newdata = test)

models <- list(
  model0 = lm(FeCl_dose ~ mean_Temp + mean_pH + mean_Conductivity + mean_Turbidity +
                Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2 +
                Turb_lag1 + Turb_lag2, data = train),

  model1 = lm(FeCl_dose ~ mean_Temp * mean_pH + mean_Turbidity + Temp_lag1  + Temp_lag2 + Turb_lag1 +
                Turb_lag2,data = train),

  model2 = lm(FeCl_dose ~ mean_Temp * mean_pH + mean_Conductivity * mean_Turbidity +
                Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2, data = train),

  model3 = lm(FeCl_dose ~ mean_Temp * mean_Conductivity + Temp_lag1 + Temp_lag2 + Cond_lag1 + Cond_lag2,                 data = train),

  model4 = lm(FeCl_dose ~ mean_Conductivity * mean_pH + mean_Temp * mean_Conductivity +
                Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2 + Cond_lag1 + Cond_lag2,
              data = train),

  model5 = lm(FeCl_dose ~ mean_Temp + mean_pH + Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2,
              data = train)
)

descriptions <- c(
  "Main effects only",
  "Temp × pH interaction",
  "Temp × pH + Cond × Turb interactions",
  "Temp × Cond interaction",
  "Cond × pH + Temp × Cond interactions",
  "Only Temp and pH"
)

performance <- data.frame(Model = character(),
                          RMSE = numeric(),
                          MAE = numeric(),
                          Description = character(),
                          stringsAsFactors = FALSE)

for(i in 1:length(models)) {
  pred <- predict(models[[i]], newdata = test)
  performance <- rbind(performance,
                       data.frame(Model = paste0("Model ", i-1),
                                  RMSE = round(Metrics::rmse(test$FeCl_dose, pred), 3),
                                  MAE  = round(Metrics::mae(test$FeCl_dose, pred), 3),
                                  Description = descriptions[i],
                                  stringsAsFactors = FALSE))

}
  performance <- performance %>% arrange(RMSE)

best_rmse <- min(performance$RMSE)

kable_table <- performance %>%
  kable("html", caption = "Model Performance Summary") %>%
  kable_styling(full_width = F, position = "center") %>%
  row_spec(which(performance$RMSE == best_rmse), background = "#cfe2f3")

kable_table

p3 <- ggplot(test, aes(x = FeCl_dose, y = pred_model3)) +
  geom_point(color = "blue") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Model 3: Predicted vs Actual FeCl_dose",
       x = "Actual FeCl_dose",
       y = "Predicted FeCl_dose") +
  theme_minimal()

p4 <- ggplot(test, aes(x = FeCl_dose, y = pred_model4)) +
  geom_point(color = "green") +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Model 4: Predicted vs Actual FeCl_dose",
       x = "Actual FeCl_dose",
       y = "Predicted FeCl_dose") +
  theme_minimal()

p3
p4

```

# Coagulant Dose Prediction Linear Regression

```{r}
knitr::include_graphics("Images/linear_model_table.png")
```

## Regression Models

- Created 6 models with different predictors and combinations of interaction terms.

- Dropped Turbidity as it was not a good predictor for FeCL dosing.

- Difficult to determine what factors of the dataset are driving the predictions.

### Performance

- Performed best with multiple interactions between Conductivity, Temp, and pH

- Still not great predictions many are overfit and underfit.

- Linear Structure doesn't capture the relations in our predictors