### 2025-12-11 Presentation DSCI445.Rmd
Final final updates

---
title: "Predicting Water Treatment Plant Chemical Doses"
author: "Ethan Schilling, Joshua Tobey, Dawson Carney, Reece Carmody"
date: "11 December 2025"
aspectratio: 169
output: 
  beamer_presentation:
    theme: "Warsaw"
    colortheme: "default"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r, include=F}
library(dplyr)
library(tidyr)
library(readr)
library(lubridate)
library(ggplot2)
library(zoo)
library(purrr)
library(knitr)
library(patchwork)
library(tidyverse)
library(tidymodels)
library(rpart.plot)
library(kableExtra)
library(vip)
library(png)
library(grid)
library(parsnip)
library(discrim)
```

```{r}
df_lag_incdec <- read.csv("RWAR_dosing_lag_incdec.csv") |>
  mutate(Timestamp = ymd(Timestamp))
df_lag_incdec$FeCl_change <- as.factor(df_lag_incdec$FeCl_change)
df_lag_incdec$Polymer_change <- as.factor(df_lag_incdec$Polymer_change)
```

# Background and Motivation

**The drinking water treatment process takes water from a river, lake, reservoir, or other source, and purifies it to have it reach drinking water standards.**

```{r}
img_array = readPNG("Images/treatmentplot1.png")
grid.raster(img_array)
```

# Background and Motivation

**An operator looks at a wide array of data and makes decisions about chemical dosing.**

**OUR GOAL: Create a model that serves as an "operator," taking raw water data and making predictions of chemical doses.**

```{r}
img_array = readPNG("Images/treatmentplot2.png")
grid.raster(img_array)
```

# Data Overview

Three years of data from 2018-2020, provided by a Colorado water treatment plant.

- **Raw Water Data**
   - pH
   - Temperature (of the water)
   - Turbidity
   - Conductivity
   - Suspended Grain Size (dropped)
   - Alkalinity (dropped)
   - Hardness (dropped)
- **Chemical Dosing Data**
   - Ferric Chloride Dose: primary
   - Cationic Polymer Dose: secondary
- Data Cleaning
   - Removed significant outliers
   - Took daily averages for raw water data
   - Merged common dates between datasets

# Exploratory Analysis

```{r, fig.width=7, fig.height=2}

plot_polymer <- ggplot(df_lag_incdec, aes(x = Timestamp,
                           y = Polymer_dose,
                           color = Polymer_dose)) +
  geom_line() + 
  scale_color_gradient(low = 'yellow', high = 'green') +
  theme_minimal() +
  ggtitle("Polymer Coagulant")

print(plot_polymer)

plot_FeCl <- ggplot(df_lag_incdec, aes(x = Timestamp,
                           y = FeCl_dose,
                           color = FeCl_dose)) +
  geom_line() + 
  scale_color_gradient(low = 'blue', high = 'orange') +
  theme_minimal() +
  ggtitle("FeCl Coagulant")

print(plot_FeCl)
```

# Exploratory Analysis

```{r, fig.width=7, fig.height=4}

df_lag_incdec_long <- df_lag_incdec |>
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "Characteristics",
    values_to = "Value"
  )

ggplot(df_lag_incdec_long, aes(x = Timestamp, y = Value)) +
  geom_point(size = 0.5, color="blue") +
  facet_wrap(~Characteristics, scales = "free_y") +
  theme_minimal() +
  labs(
    x = "Date",
    y = "Value",
    title = "Raw Water Characteristics"
  )

```

# Predictor Selection

Based on preliminary analyses, we will focus on the following dose predictors:

- **pH** (strongest relationship)
- **Temperature** (strongest relationship)
- **Conductivity** (some noise, weaker)
- **Turbidity** (lots of noise, weaker)

Outcome: **FeCl Dose**. (Polymer dose is not changed significantly in plants.)

# Added Variables

Introduced **lagged variables** like in the Stock dataset from the homework.

- For each of the main four variables:
   - Introduce two lagged variables
   - First represents the day before
   - Second represents two days before

Introduced a **categorical predictor** indicating whether the dose increased, decreased, or stayed the same compared to the day before.

# Summary of Modeling

**Priorities: Interpretability, Capturing of Nonlinear Behavior**

- Training and Testing
   - Training 2018-2019, Testing 2020
   - Approximately 70/30 split
- **Coagulant Dose Prediction**
   - Linear Regression Models
   - Tree-Based Models
- **Coagulant Dose Change Prediction**
   - LDA, QDA Classification
   - Multinomial Classification
   - Random Forest Classification

# Coagulant Dose Prediction

**GOAL: Predict coagulant doses based on raw water characteristics.**

```{r, fig.width=7, fig.height=3}

df_lag_incdec_long <- df_lag_incdec |>
  pivot_longer(
    cols = starts_with("mean_"),
    names_to = "Characteristics",
    values_to = "Value"
  )

ggplot(df_lag_incdec_long, aes(x = Timestamp, y = Value, color = FeCl_dose)) +
  geom_point(size = 0.7) +
  facet_wrap(~Characteristics, scales = "free_y") +
  scale_color_viridis_c(option = "plasma") +
  theme_minimal() +
  labs(
    x = "Date",
    y = "Value",
    color = "Ferric Chloride (mg/L)",
    title = "Raw Water Characteristics Over Time Colored by Dosing"
  )

```

# Coagulant Dose Prediction: Linear Regression

- Created 6 models with different predictors and combinations of interaction terms

- Used lag variables for temp, pH, conductivity, turbidity

- Dropped Turbidity as it was not a good predictor for FeCL dosing

- Difficult to determine what factors of the dataset are driving the predictions

- RMSE and MAE vary widely depending which predictors are selected and the interactions

# Coagulant Dose Prediction: Linear Regression

```{r, out.width="80%", out.height="88%"}
knitr::include_graphics("Images/linear_model_table.png")
```

# Coagulant Dose Prediction: Decision Tree

```{r, fig.width=5, fig.height=2.5}
set.seed(445)

# Decision Tree

# For now just predicting numerical amounts, not classifying
data = read.csv("RWAR_dosing_lag_incdec.csv")
data = as.data.frame(data)
data$FeCl_change = NULL
data$Polymer_change = NULL

# Set up training & testing data

# Can change these if you want to modify the tree/data
tree_depth = 5 # tree depth
min_n = 50 # minimum observations needed to further split

train_data = data %>%
  filter(year(Timestamp) <= 2019)
test_data = data %>%
  filter(year(Timestamp) >= 2020)

# Create the tree

recipe = recipe(FeCl_dose ~ ., data = train_data) %>% # Change target variable here
  step_rm(Timestamp) %>%
  step_rm(Polymer_dose) # Remove the other target variable here

tree_spec = decision_tree(tree_depth = tree_depth, min_n = min_n) %>%
  set_engine("rpart") %>%
  set_mode("regression")

workflow = workflow() %>%
  add_recipe(recipe) %>%
  add_model(tree_spec)

fit_tree = workflow %>%
  fit(data = train_data)

# Plot the tree

fit_tree %>%
  extract_fit_engine() %>%
  rpart.plot(roundint = FALSE)

# Assess accuracy

predictions = fit_tree %>%
  predict(test_data) %>%
  bind_cols(test_data)

actual = predictions$`FeCl_dose` # Change target variable here
predicted = predictions$.pred

rmse = sqrt(mean((actual - predicted)^2))
r_squared = cor(actual, predicted)^2
mae = mean(abs(actual-predicted))

metrics = data.frame(
  Metric = c("RMSE", "R-Squared", "MAE"),
  Value  = c(rmse, r_squared, mae)
)

kable(metrics, booktabs = TRUE, digits = 3) %>%
  kable_styling(position = "center", font_size = 12)

```

# Coagulant Dose Prediction: Random Forest

```{r, fig.width=5, fig.height=2.1}
set.seed(445)

# Random Forest

# For now just predicting numerical amounts, not classifying
data = read.csv("RWAR_dosing_lag_incdec.csv")
data = as.data.frame(data)
data$FeCl_change = NULL
data$Polymer_change = NULL

# Set up training & testing data

# Can change these if you want to modify the tree/data
trees = 500 # number of trees in the forest
min_n = 50 # minimum observations needed to further split

train_data = data %>%
  filter(year(Timestamp) <= 2019)
test_data = data %>%
  filter(year(Timestamp) >= 2020)

# Create the tree

recipe = recipe(FeCl_dose ~ ., data = train_data) %>% # Change target variable here
  step_rm(Timestamp) %>%
  step_rm(Polymer_dose) # Remove the other target variable here

rf_spec = rand_forest(trees = trees, min_n = min_n) %>%
  set_engine("ranger", importance = "impurity") %>%
  set_mode("regression")

workflow = workflow() %>%
  add_recipe(recipe) %>%
  add_model(rf_spec)

fit_rf = workflow %>%
  fit(data = train_data)

# Assess accuracy

predictions = fit_rf %>%
  predict(test_data) %>%
  bind_cols(test_data)

actual = predictions$`FeCl_dose` # Change target variable here
predicted = predictions$.pred

# Metrics and plots

model_metrics = predictions %>%
  metrics(truth = FeCl_dose, estimate = .pred) # Change target variable here
rmse = model_metrics$.estimate[1]
r_squared = model_metrics$.estimate[2]
mae = model_metrics$.estimate[3]

metrics = data.frame(
  Metric = c("RMSE", "R-Squared", "MAE"),
  Value  = c(rmse, r_squared, mae)
)

vip(fit_rf, num_features = 20) +
  geom_col(color = "black", fill = "coral") +
  labs(
    title = "Feature Importance (Random Forest)",
    x = "Importance",
    y = "Feature"
  ) +
  theme_minimal(base_size = 9)

kable(metrics, booktabs = TRUE, digits = 3) %>%
  kable_styling(position = "center", font_size = 12)

```

# Coagulant Dose Change Prediction

**GOAL: Predict whether coagulant dose increased, decreased, or stayed the same based on raw water characteristics.**

```{r, fig.width=7, fig.height=3}

ggplot(df_lag_incdec_long, aes(x = Timestamp, y = Value, color = FeCl_change)) +
  geom_point(size = 0.7) +
  facet_wrap(~Characteristics, scales = "free_y") +
  theme_minimal() +
  labs(
    x = "Date",
    y = "Value",
    color = "FeCl_change",
    shape = "FeCl_change",
    title = "Raw Water Characteristics Colored by Dose Change"
  )

```

# Coagulant Dose Change Prediction: LDA

- Experimented with predictor sets and found that pH and temperature (plus the associated lagged variables) were the main effective predictors.
- Turbidity and conductivity did not improve prediction.

LDA ACCURACY: 43.7%

```{r}
set.seed(445)

## TRAIN/TEST DATA (2018-2019 and 2020)
df_full <- df_lag_incdec |> select(-c(FeCl_dose, Polymer_dose))
df_train <- df_full |> filter(year(Timestamp) <= 2019)
df_test  <- df_full |> filter(year(Timestamp) == 2020)

## LDA
lda_spec <- discrim_linear()
lda_fit1 <- lda_spec |>
  fit(FeCl_change ~ ., data = df_train |> 
        select(-c(Polymer_change, Timestamp)))
lda_fit2 <- lda_spec |>
  fit(FeCl_change ~ mean_pH + mean_Temp + pH_lag1 + pH_lag2 + Temp_lag1 + Temp_lag2, data = df_train)

# select LDA to take metrics of
lda_fit <- lda_fit2

# performance metrics
lda_fit_disp <- lda_fit |> extract_fit_engine()
lda_conf <- lda_fit |> 
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)
lda_acc <- lda_fit |>
  augment(new_data = df_test) |>
  accuracy(truth = FeCl_change, estimate = .pred_class)

lda_table <- lda_conf$table |> 
  as.data.frame() |>
  rename(Predicted = Prediction, True = Truth, Count = Freq)
lda_table |>
  tidyr::pivot_wider(names_from = Predicted, values_from = Count) |>
  knitr::kable("markdown", caption = "LDA Confusion Matrix")

## Notes
# 39% test accuracy on all predictors
# 41% test accuracy on temp and pH

```

# Coagulant Dose Change Prediction: QDA

- Experimented with predictor sets and found that pH and temperature (plus the associated lagged variables) were the main effective predictors.
- Turbidity and conductivity did not improve prediction.

QDA ACCURACY: 35.7%

```{r}
set.seed(445)

## QDA
qda_spec <- discrim_quad()
qda_fit1 <- qda_spec |>
  fit(FeCl_change ~ ., data = df_train |> 
        select(-c(Polymer_change, Timestamp)))
qda_fit2 <- qda_spec |>
  fit(FeCl_change ~ mean_pH + mean_Temp + pH_lag1 + pH_lag2 + Temp_lag1 + Temp_lag2, data = df_train)

# select QDA to take metrics of
qda_fit <- qda_fit2

# performance metrics
qda_fit_disp <- qda_fit |> extract_fit_engine()
qda_conf <- qda_fit |> 
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)
qda_acc <- qda_fit |>
  augment(new_data = df_test) |>
  accuracy(truth = FeCl_change, estimate = .pred_class)

qda_table <- qda_conf$table |> 
  as.data.frame() |>
  rename(Predicted = Prediction, True = Truth, Count = Freq)
qda_table |>
  tidyr::pivot_wider(names_from = Predicted, values_from = Count) |>
  knitr::kable("markdown", caption = "QDA Confusion Matrix")

## Notes
# 39.5% test accuracy on all predictors
# 43.7% test accuracy on temp and pH

```

# Coagulant Dose Change Prediction: Multinomial Classification

- Experimented with predictor sets and found that pH and temperature (plus the associated lagged variables) were the main effective predictors.
- Turbidity and conductivity did not improve prediction.

MULTINOMIAL ACCURACY: 46.3%

```{r}
## MULTINOMIAL REGRESSION
multi_model <- multinom_reg() |>
  set_engine("nnet") |>
  set_mode("classification")

multi_recipe1 <- recipe(FeCl_change ~ mean_Temp + Temp_lag1 + Temp_lag2 + mean_pH + pH_lag1 + pH_lag2 + mean_Conductivity + Cond_lag1 + Cond_lag2, data = df_train) |>
  step_normalize(all_numeric_predictors()) |>
  step_dummy(all_nominal_predictors())

multi_recipe <- multi_recipe1

multi_workflow <- workflow() |>
  add_recipe(multi_recipe) |>
  add_model(multi_model)

multi_fit <- fit(multi_workflow, data = df_train)

df_pred_class <- predict(multi_fit, new_data = df_test, type = "class") |>
  bind_cols(df_test |> select(FeCl_change))
df_pred_probs <- predict(multi_fit, new_data = df_test, type = "prob") |>
  bind_cols(df_test |> select(FeCl_change))

# results
multi_acc <- metrics(df_pred_class, truth = FeCl_change, estimate = .pred_class) |> 
  filter(.metric == "accuracy")
multi_conf <- conf_mat(df_pred_class, truth = FeCl_change, estimate = .pred_class)
multi_fit_disp <- multi_fit |> extract_fit_engine()

multi_table <- multi_conf$table |> 
  as.data.frame() |>
  rename(Predicted = Prediction, True = Truth, Count = Freq)
multi_table |>
  tidyr::pivot_wider(names_from = Predicted, values_from = Count) |>
  knitr::kable("markdown", caption = "Multinomial Regression Confusion Matrix")

## NOTES
# turbidity is highly variable and does not help

```

# Coagulant Dose Change Prediction: Random Forest

- Experimented with predictor sets and found that pH and temperature (plus the associated lagged variables) were the main effective predictors.
- Turbidity and conductivity did not improve prediction.

RANDOM FOREST ACCURACY: 39.2%

```{r}

set.seed(445)

## RANDOM FOREST
rf_spec <- rand_forest(mtry = sqrt(.cols()), trees=500, min_n=1) |>
  set_engine("randomForest", importance = TRUE) |>
  set_mode("classification")

# Different model fits
rf_fit1 <- rf_spec |>
  fit(FeCl_change ~ ., data = df_train |> 
        select(-c(Polymer_change, Timestamp)))
rf_fit2 <- rf_spec |>
  fit(FeCl_change ~ mean_Temp + mean_pH + Temp_lag1 + Temp_lag2 + pH_lag1 + pH_lag2, 
      data = df_train)
rf_fit3 <- rf_spec |>
  fit(FeCl_change ~ mean_pH + mean_Temp, 
      data = df_train)

# Model fit to take performance metrics of:
rf_fit <- rf_fit2

# Most important predictors
#vip(rf_fit)

## results
rf_acc <- rf_fit |>
  augment(new_data = df_test) |>
  accuracy(truth = FeCl_change, estimate = .pred_class)
rf_conf <- rf_fit |>
  augment(new_data = df_test) |>
  conf_mat(truth = FeCl_change, estimate = .pred_class)

rf_table <- rf_conf$table |> 
  as.data.frame() |>
  rename(Predicted = Prediction, True = Truth, Count = Freq)
rf_table |>
  tidyr::pivot_wider(names_from = Predicted, values_from = Count) |>
  knitr::kable("markdown", caption = "Random Forest Confusion Matrix")

## Notes
# 34% accuracy for all predictors
# 39% accuracy for temp + pH + temp/pH lags
# 46.6% accuracy for just temperature and pH

```

# Summary of Dose Prediction Models

```{r, fig.height=3.8, fig.width=7}

# - The model with pH & temperature interactions with conductivity performed the best out of models tested.
# - Tree-based models performed the worst, likely due to overfitting.
# - All models diverged from the actual dose by an average of at least 1.4 mg/L.

dosepred_df <- data.frame(Model = c("Random Forest", "Decision Tree", "Cond-pH + Cond-Temp", "Conductivity-Temp", "Temp and pH"),
                          RMSE = c(2.811, 2.756, 1.672, 1.891, 2.350),
                          MAE = c(2.283, 2.484, 1.421, 1.516, 1.941))

# convert to long format
dosepred_long <- dosepred_df |>
  arrange(RMSE) |>
  mutate(Model = factor(Model, levels = unique(Model))) |>
  pivot_longer(
    cols = c(RMSE, MAE),
    names_to = "Metric",
    values_to = "Value"
  )

# Plot generation
ggplot(dosepred_long, aes(x = Model, y = Value, fill = Metric)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8), width = 0.7) +
  geom_text(
    aes(label = sprintf("%.3f", Value)), 
    position = position_dodge(width = 0.8),
    vjust = -0.5, 
    size = 2
  ) +
  labs(
    title = "Dose Prediction Model Comparison (RMSE and MAE)",
    x = "Model",
    y = "Error Value"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

# Summary of Dose Change Prediction Models

Overall ineffective, and prone to overfitting (especially random forest).  
Model predicting no dose change would be more accurate.

```{r, fig.width=7, fig.height=3}

dosechangepred_df <- data.frame(
  Model = c("LDA", "QDA", "Multinomial", "Random Forest"),
  Accuracy = c(lda_acc$.estimate, qda_acc$.estimate, 
               multi_acc$.estimate, rf_acc$.estimate)
) |> arrange(Accuracy)

ggplot(dosechangepred_df, aes(x = reorder(Model,-Accuracy), y = Accuracy)) +
  geom_col(fill = "steelblue") +
  geom_text(aes(label = scales::percent(Accuracy, accuracy = 0.1)),
            vjust = -0.3, size = 4) +
  scale_y_continuous(labels = scales::percent_format(accuracy = 1),
                     limits = c(0, max(dosechangepred_df$Accuracy) * 1.1)) +
  labs(x = "Model", y = "Accuracy") +
  theme_minimal()

```

# Conclusion and Next Steps

- Predicting chemical doses based on these raw water characteristics proved to be ineffective
- Most likely reasons:
   - Limited size of dataset
   - Limited predictor set
   - Operator-dependent
   - Complexity of chemical relationships
   - Other factors (example: reservoir turnover)
- Operators make decisions based on variety of other factors: chemical waste production, filter performance, lab tests, etc.
- Incorporate more varieties and time ranges of data in future modeling efforts